### 2020-06-12 Paper.Rmd
Correction of parameters, and first draft of materials and methods

---
title: "Draft 1"
author: "Ana Karina"
date: "4/7/2020"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

# Risk of infection from touching surfaces contaminated with SARS-CoV-2
## Abstract
## Introduction

close contact route (large droplet), airborne (small droplets), fomites.
An investigation of a nosocomial outbreak of COVID-19 reported at Augustine’s Hospital in Durban, South Africa suggest that indirect contact through health care workers and fomites were the predominant transmission pathways @Lessells2020.

*Risk assessment models*
Mathematical models have been used to understand SARS-CoV and MERS-CoV @Adhikari2019 @Lei2017.
Let et al @Lei2017  developed a model to understand the relative contribution of different transmission routes (airborne, close contact, and fomites) on the risk of infection for influenza A H1N1, SARS-CoV, and norovirus. The simulation results suggest that around 50% of the CoV infections were due to indirect (fomite mediated) transmission. A QMRA of MERS-CoV was developed by Adhikari et all @Adhikari2019 to evaluate the risk for hospital transmission of coronavirus. They only considered the risk of infection due from droplets and aerosolized particles, since that is the main transmission route. Most of the uncertainty associated with the model was due to the concentration of virus in the saliva. The researchers evaluated possible interventions and showed that the use of mask contributed significantly to the reduction of the risk (>90%), as well as increasing air exchange (up to 58% risk reduction).

extensive environmental contamination has been found in settings with COVID-19 patients @Ong2020 @Ye2020, coronaviruses persist on inanimate surface for extended periods of time @Kampf2020

If no countermeasures are taken, viruses can spread through the space populated by the human host at a high speed, as we have observed @BarOn2020.

Surfaces play a significant role in the spread of respiratory viral diseases @Boone2007

Potential role of aerosol transmission of SARS-CoV-2, sampling the air @Liu2020

The CDC and WHO have highlighted two main transmission routes: direct transmission or person-to-person transmission (respiratory droplets produced while talking, coughing, or sneezing), and indirect transmission or surface-mediated transmission (touching a contaminated surface followed by touching mucose membranes in the face such as mouth or nose) @WHO2020 @CDC2020.

Although indirect transmission through fomite mediated contacts is not the primary transmission route, it can lead to infection @Chen2004. A study of an outbreak of SARS CoV in Taiwan showed that, out of the 232 SARS-CoV patients in the hospital, 31 patients (13 %)  were hospital-acquired transmissions. Out of the 31 cases, most of the transmission occurred by direct contact with a known SARS patient, but 6/31 (~20%) of the transmissions could not be traced to direct contact with SARS patients. Sampling of surfaces for the RNA of coronavirus showed that surfaces such as drinking water buttons, beds and chairs were contaminated with coronavirus. Thus, the researchers postulated that indirect contact could have been the source of transmission @Chen2004.

Viruses are known to transfer from contaminated surfaces to the skin upon contact @Julian2010 @Lopez2013 @Bidawid2004 @Rusin2002 @Ansari1991 and from the skin to the mouth and saliva of individuals @Pitol2017 @Rusin2002. People touch their faces frequently @Nicas2008 @Kwok2015. Therefore, touching contaminated surfaces offers a risk of exposure. Despite its importance, there is not a lot of information regarding the magnitude of risk associated with touching contaminated surfaces @BarOn2020.

Hand hygiene interventions and surface disinfection have been highlighted as one of the primary strategies to control transmission @Mcclellan2020. Studies with coronaviruses have shown than hygiene practices such as handwashing can reduce the risk of coronavirus infection @Beale2020. But there is no conclusive evidence of the effectiveness of hand-washing to reduce SARS-CoV-2 transmission @Yang2020.

Relevance of fomites in the context of SARS-CoV-2 (Stanford review) @Castano2020

## Materials and Methods 
### Hazard Identification
Whenever an infected individual coughs, sneezes or speaks, droplets containing pathogens make their way to the proximal surfaces [@Bourouiba2014]. It has been shown that SARS-CoV-2 can persist on surfaces like plastic, metal, or glass at room temperature for hours to days [@VanDoremalen2020]. Additionally, viruses are known to transfer from surfaces to the hands upon contact [@Julian2010; @Lopez2013; @Bidawid2004; @Rusin2002; @Ansari1991], as well as from hands to the lips and saliva of individuals [@Rusin2002; @Pitol2017]. Given the fact that people frequently touch their faces and the mucous membranes within the face (nose, eyes and mouth) [@Kwok2015; @Nicas2008], we have identified the indirect (surface-mediated) transfer of SARS-CoV-2 as a relevant transmission pathway. This risk assessment is designed to explore the risks of infection associated with the surface-mediated transmission of SARS-CoV-2.

### Exposure Assessment 
Only one potential exposure pathway was analysed: the indirect transmission of viruses through contaminated surfaces. Two different scenarios where modelled: single touch (frequently touched buttons) and multiple touches (ATM or cash machine).

*Scenario 1. Single touch buttons.* Frequently touched buttons such as traffic light buttons, train buttons, and elevator buttons are of risk concern due to their frequent use by multiple people. In this scenario, we modelled the risk of touching a contaminated button followed by touching mucous membranes on the face. The material of the button was assumed to be either plastic or steel. The sequence of events leading to the viral inoculation of the susceptible individual was considered to be the following: 1) contamination event (infected person coughing directly on the surface), 2) virus decay on the surface, 3) virus transfer from the surface to the finger, 4) virus transfer from the finger to the mucous membranes in the face (eyes, mouth, nose).

*Scenario 2. Multiple touches.* ATM/cash machines are another example of surfaces that are frequently visited by individuals and are not frequently cleaned, as they are located in public spaces. For this risk assessment model, the ATM was assumed to be made of steel. The sequence of events leading to virus inoculation was considered to be the following: 1) contamination event (infected person coughing directly on the surface), 2) virus decay on the surface, 3) virus transfer from the surface to the finger and from finger to surface through multiple contacts, 4) virus transfer from the finger to the mucous membranes in the face (eyes, mouth, nose).

####  Concentration and transfer of SARS-CoV-2
Viral loads in the saliva or sputum of symptomatic COVID-19 patients within the first 14 days of symptom onset were used as model input. The virus titers used in the risk assessment (Table 1) were measured using quantitative reverse-transcription polymerase chain reaction (RT-qPCR)  [@Wolfel2020; @Pan2020; @Kim2020; @To2020]. The RT-qPCR method estimates the number of copies of viral RNA in the clinical samples. Nevertheless, the presence of coronavirus RNA does not imply the presence of infectious viruses. In risk assessments where no data on the concentration of infectious viruses is available, it is commonly accepted to use a gene copies to infectivity ratio to convert the concentration measured in gene copies to the concentration of infective viruses (REF). The gene copy number to infectivity ratio is, to our knowledge, not available for SARS-CoV-2. In the absence of SARS-CoV-2 data, we used the relationship reported for seasonal A(H1N1), influenza A(H3N2), and influenza B @Ip2016 (Table 1). Both, influenza viruses and coronaviruses, are enveloped, RNA viruses with similar sizes.

The concentration of viruses in the surface after inoculation (through cough) was estimated using the following equation:

\begin{equation}
C(0)=\frac { C_{sp}V_{s}}{\frac {4 \pi x_{i}^{2}}{a}}
\end{equation}

where $C(0)$ [virus/cm^2^] is the concentration of virus in the surface, $V_{s}$ [mL/cough] is the volume of saliva expelled per cough, $C_{sp}$ [virus/mL] is the concentration of infective viruses in the saliva or sputum, $x_i$[cm] is the inoculation distance or the distance between the surface and the infected individual at the time of the cough for activity $i$, and $a$ is the XXX. This equation assumes that the droplets are expelled in a conical shape and a proportion of the spheric part of the cone is in contact with the surface. After surface inoculation, the concentration of the virus on the surface at time = $t$ was estimated using the following equation:

\begin{equation}
C(t)= C(0)  e^{-nt}
\end{equation}

where $C(t)$[virus/cm^2^] is the concentration of infective viruses at time $t$, $C(0)$[virus/cm^2^] is the initial concentration of virus in the surface, $t$[min] is the time lapsed after surface inoculation, and $n$[m^{-1}^] is the decay rate of the virus in the selected surface (metal or plastic). The number of viruses transferred from the surface to the finger $N_f$ [virus] was estimated using the following equation:

\begin{equation}
N_f = C(t)  TE_{sh} A_f
\end{equation}

where $TE_{sh}$[%] is the transfer efficiency of viruses between surfaces and hands, $C(t)$ [virus/cm^2^] is the concentration of virus in the surface and $A_f$[cm^2^] is the area of the finger. For multiple touch scenarios the number of virus in the finger multiple contacts was calculated as follows:

The dose of viruses that entered the susceptible individual through facial membranes, $D$ [virus], was estimated as follows:

\begin{equation}
D = N_f FSA_{fm} TE_{hm}
\end{equation}

where $N_f$[virus] is the number of virus in the finger, $FSA_{fm}$ [%] is the fractional surface area between the finger and the mucous membranes, and $TE_{hm}$ [%] is the transfer efficiency of the virus from the hand to the mucous membranes.

#### Dose-response
In the case of highly virulent pathogens, where no prophylaxis is available, it is common practice to use data of animal trials to generate dose-response models. The dose-response relationship for SARS was obtained from the QMRA wiki <http://qmrawiki.org/experiments/sars>. This model is based on two studies [@DeAlbuquerque2006; @DeDiego2008] of SARS-CoV and Murine hepatitis virus (MHV-1) infection in mice. MHV is a commonly used surrogate for human coronaviruses.

The risk of infection was estimated using the following exponential dose-response model:

\begin{equation}
P_{inf}=1 - e^{-kD}
\end{equation}

where $P_{inf}$ [%] is the probability of infection, $k$ [-] is the infectivity parameter of the exponential model and $D$ [virus] is the dose of virus that entered the individual through mucous membranes.

#### Model parameters used in the risk assessment 
| Parameter | Units | Description | Input values / Equation | Distribution | Reference and comments |  
|:--:|:--:|:--:|:--:|:--:|:--:|
| $C_{spGC}$ | gene copies/mL | Concentration of SARS-CoV-2 in sputum or saliva | Log~10~ gene copies/mL (data set) | ReSample | [@Wolfel2020; @Pan2020; @Kim2020; @To2020] RT-qPCR data from 9, 2, 2, and 23 patients with COVID-19|
| $GC:TCID_{50}$ | - | Genome copies to infectious virus conversion factor | 10^2^-10^3^ | Uniform | [@Ip2016] Mean RT-qPCR to TCID~50~ ratio of 708, 547, and 185 gene copies per TCID~50~ for seasonal A(H1N1), A(H3N2), and influenza B, respectively |
| $C_{sp}$ | TCID~50~/mL |Concentration of infective SARS-CoV-2 in sample | $\frac {C_{spGC}}{GC:TCID_{50}}$ | - | calculated|
| $V_s$ | mL | Volume of saliva expelled per cough | 0.0484-0.0396 | uniform | [@Nicas2009] volume = 0.044mL, [@Adhikari2019] uniform $\pm$ 10%|
|$t_{ATM}$ | min |Time spent on the ATM | 0.75(0.38) | Normal | This study. Otained from 20 hours of observational data |
|$t_{btw-ATM}$ | min |Time between visits | 8.4(7.1) | Normal | This study. Otained from 20 hours of observational data |
|$x_{ATM}$| cm | ATM inoculation distance | 40 | Point value | This study. Assumed from the observational data  |
|$x_{btn}$| cm | Button inoculation distance | 50 | Point value | Assumed |
|$t_{0.5-stl}$| min | Half life of CoV-2 in steel | 338(35) | Normal | @VanDoremalen2020, SARS-CoV-2 infectivity assays, 40% RH and 21-23°C |
|$t_{0.5-pl}$| min | Half life of CoV-2 in plastic | 409(39) | Normal | @VanDoremalen2020, SARS-CoV-2 infectivity assays, 40% RH and 21-23°C |
| $n_x$| min^-1^ |Decay rate in surface $x$ (steel or plastic) | $\frac {ln(2)}{t_{0.5}}$ | - | calculated |
| $TE_{sh-stl}$ | % | Transfer efficiency of virus between steel to hand | 37.4(16) | Normal | @Lopez2013, transfer efficiency of viruses (MS2) between surface and skin at 40-65% humidity | 
| $TE_{sh-pl}$ | % | Transfer efficiency of virus between plastic to hand | 79.5(21.2) | Normal | @Lopez2013, transfer efficiency of viruses (MS2) between surface and skin at 40-65% humidity | 
| $TE_{hm}$ | - | Transfer efficiency from hand to mucouse membranes | 0.201(0.063) | Normal | @Pitol2017, transfer efficiency of viruses (MS2) from hand to saliva | 
| $A_{f}$ | cm^2^ | Finger surface area | 3.8-6.6 | Uniform | Data from @Chabrelie2018 @ @Peters2009a @Sahmel2015 | 
| $FSA_{fm}$ | - | Fractional surface area between finger and mucous membranes | 0.5-0.8 | Uniform | Assumed | 
| $k$ | PFU^-1^ | Parameter of the exponential dose-response| 0.00107, 0.00135, 0.00680 | Triangular | Data obtained from QMRAwiki, based on 2 studies @DeAlbuquerque2006, @DeDiego2008, using the from 0.5th, 50th, and 99.5th percentiles as min, mode, and max |
|:--:|:--:|:--:|:--:|:--:|:--:|
| N~cgh/min~ |min^-1^ | Number of coughs per minute  | 0.57(1)| Normal |@Leung2020, based on data of 17 individuals with seasonal CoV|

### Risk Characterization 
Monte Carlo simulation method is used to incorporate uncertainty and variability of the input parameters in the risk characterization.

### Sensitivity analysis

A sensitivity analysis was performed to investigate hot the variabilty and uncertainty of the parameters in the model onfluenced the estimated risks. The sensitivity was estimated using Spearman correlation coefficients between the inputs and outputs of the model.

## Results and Discussion
```{r dose-response, include=FALSE}

```

### Factors influencing fomite mediate virus risk
There are number of factors influencing indirect risk of SARS-CoV-2 transmission @Moriyama2020, some of these factors are highly correlated with each other. Seasonal and environmental factors such as temperature, humidity and sunlight are examples of them. Relative humidity (RH) influences the size of the droplets expelled in the cough, virus survival on surfaces @Casanova2010, and virus transfer between hands and surfaces @Lopez2013. Lopez et al showed that XXX. This risk assessment was performed using the transfer efficiency of viruses at a relative humidity of (X-X). If we were to exchange this variable for the percentage transfer of a low relative humidity setting we will have a lower inactivation, as the lower the RH, the higher the survival @Casanova2010, but we would have a significantly higher transfer of viruses between surfaces and skin (i.e., XX ). Overall, this change would predict a XX% decrease in the indirect risk of transmission with a dicrease in RH. Temperature and sunlight are known to correlate negatively to virus survival in surfaces @Chin2020, @Casanova2010. In addition to environmental factors, behavioural factors such as hygiene practices and social behaviour are known to influence fomite mediated transmission (REF).

One important parameter not considered in the model is the susceptible individual immune system and how the environmental factors influence it. The immune system plays an important role in the probability of infection @Moriyama2020.

Surfaces: SARS-CoV-2 is more stable in smooth surfaces @Chin2020.

Surfaces are easy to clean. SARS-CoV-2 is susceptible to disintectants including household belach and ethanol @Chin2020 @Castano2020. Although disinfection can be achieved easily using a variaty of disinfectants, public surfaces that are frequently contacted but not frequently cleanned present a challenge. Strategies to minimize the contact with surfaces such as trafic light buttons could therefore be beneficial to reduce the risk of expossure.

### Sensitivity analysis
The factors that influence the most the risk of ilness where the same in all the scenations: 1) concentration of virus in the saliva, 2) ..., 3)...

“A dX change in input results in a dY change in outcome”

Since no randomized control trials will be runned and there are limits to the conclusions we can reach using epidemiological data, risk assesments provide an attempt to understand the risk associated with different activities. The results of this paper could inform policy. For example, trafick lights in the crossing stations could clausured momentaniely, the same than train buttons.

The dependence of temperature and humidity in the transfer rate of coronavirus is unclear. If temperature and humidity may play a role in non-dominant transmission pathways, such as indirect transfer, it would be hard to observe epidemiological studies. It is known that the survival of viruses on surfaces is dependent on environmental factors such as temperature and humidity (REF). Similarly, the transfer rate of viruses varies with humidity (REF). Nevertheless, if only a small fraction of the transmission occurs through indirect transfer we wont see it.

The sensitivity analysis shows that XXX parapeters are responsible for XX % of the risk variation of the model. Collecting more data on these parameters could reduce uncertainty.

It is important to notice that the effectiveness of interventions such as hand washing depends not only on many factors which include: the hand-wasihng technique, the suceptibility of the virus to phisical and chemical desinfection, and the relative contribution of the indirect transfer of viruses to the overal infection risk. As it has been pointed out, thand hygiene should be promoted but the risk of overemphasizing it without more intormation could turn out to be negative (Yang, 2020). ...  ("risk compensation"?).

Seasonal factors such as temperature, humidity and sunlight are known to influence the spread of respiratory viruses such as influenza (REF). The role that those factos have on the spread of SARS-CoV-2 remains unknown (REF), althouh it is know that the temperature and humidity influences the viability of the virus (REF), as well as SARS-CoV-2 survive longer at lower temperature (REF).

It has been shown that seasonal environmental factors such as temperature and humidity influence respiratory virus survival and transmission @Moriyama2020.

Humidity influences the size of the droplets generated by the infected individual, the size of the droplets influence both the lifetime of the virus in the environment. 
The deposition of droplets onto surfaces depends on the particle size.

Nevertheless, our results suggest that fomite mediated transmission will be .. influenced by those factors. The extend that this influence will have in the droplet dynanimcs by influencing evaporation rate of the droplets (REF). transmission of CoV will depend on the relative contribution of indirect transmission pathway.

### Limitations

*Dose Response* 
The use of animals models in QMRA is a generally accepted @Haas2014, @Haas2015. 
It has been shown that different viruses of the family of coronaviruses have significantly different dose-response relationships, the difference in the ID~50~ or dose require to observe a 50% response can be orders of magnitude different beween different coroaviruses [@Watanabe2010]. Therefore, a dose-response relationship developed specifically for SARS-CoV-2 would increase the accuracy of the results of the present and future risk assessments. The dose response relationship used herein predicts a 50% response when the dose is 282 PFU. 
*xx*

*Virus inoculation* 
More detailed information regarding the number and size of particles expelled per couugh could be used to increase accuracy of risk assessments @Lee2019. This QMRA had big assumptions regarding the inoculation of viruses on the surfaces. Regardless of the assumptions, we expect the results not to change significantly with an increase in accurace of virus inoculation calculation, as we don't expect it to be wrong more than one order of magnitude, as compared with the variation expected in the concentration of viruses on the saliva for symptomatic patients within the first two weeks of symptom onsent, which goes from 0 to ~10^X gene copies/mL. 
*xx*

*Virus concentration on infected individuals*

There is a lack of data regarding the viral loads of infectious viruses in the saliva of infected individuals. Therefore, we translated the available concentration data, obtained using quantitative reverse-transcription polymerase chain reaction (RT-qPCR), which gives a measure of the number of copies present a viral RNA sample to infectious virus data.

Care most been taken when using viral RNA data to drive conclusions about infective viruses released by the patient, as PCR data does not distinguish infectios and non-infectious virus @Atkinson2020. The use of conversion factor from PCR data has been used in the micribial risk assessment context, due to the limit availability of infetivity data (REF). In the absence of GC:PFU data on SARS-CoV-2, we used the relationship reported for Pandemic influenza virus H1N1, as both are enveloped, RNA viruses with similar sizes. 
Currently we have liminted data on the relationship between copy numbers and infectivity, which limits the extend to which the risk assessment can be used. However, modeling approaches such as the QMRA can help elucidate the contribution of different exposure routes to the disease transmission and therefore help the development of guidelines. 
*xx*

*Transmission routes*
We recomend to further develop risk assessment models, incorporating other transmission routes, to understand the relative countributionof each route. 
*x*

Most of the data used in the models are coming from laboratory experiments. The relevance of the laboratory conditions to real world conditions depend on the parameters.

*Conclusion*
Even though this risk assessment model relies on many assumptions, some with high uncertainty around them, and the risk caucluation is hihgly simplidied, this model can: provide insigts into the possible effects that different interventions can have in reducing the spread of SARS-CoV-2, serve as a first step in the pathway of building more complex models that allow exploring the relative contribution that different transmission routes could have in spreddin the disease.

about the paper, we can say something like "Although the relative importance of each transmission route remains poorly understood, there has been a growing consensus that contaminated fomites play a critical role in the spread of viruses." @Boone2007

The relative contribution of surface-mediated transmission as compared with other transmission routes such as direct transmission, air, or thorugh the feca-oral route remains to be addressed. QMRA can be used to incorporate different transmission routes and answer this question.

QMRA can be used to assess the significance of surface mediated transmission of SARS-CoV-2 [@Haas2020], as it has been used in the past to assess the risk of other transmission routes for MERS-CoV [@Adhikari2019]. Although not perfect, a risk assessment framework could be use to conduct a hazzard assessment. Since offices will need to re-open https://www.osha.gov/shpguidelines/hazard-Identification.html.

Oportunities for managing the risks associated with people's interaction with frequently touched surfaces include contact reduction at a population level, for example reduce the need for contacting trafic ligth buttons by setting them into automatic mode. Since surface cleanning of public spaces is not feasible with the frequency needed to significanlty reduce risks, signs that remind people to clean their hands after contact could potentially be encoraged (REF).

## Supporting Information

## Author Information

## Aknowledgements

## References