### 2020-09-11 ES&T paper/Supporting_Information.tex

\documentclass[journal=jacsat,manuscript=article]{achemso}
\usepackage[version=3]{mhchem}
\usepackage{amsmath}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\author{Ana K. Pitol}
\affiliation{Department of Civil and Environmental Engineering, Imperial College
London, United Kingdom}
\email{a.pitol-garcia@imperial.ac.uk}
\author{Timothy R. Julian}
\affiliation{Department of Environmental Microbiology, Eawag, Dubendorf, Switzerland}
\alsoaffiliation{Department of Epidemiology and Public Health, Swiss Tropical and Public
Health Institute, Basel, Switzerland}
\email{tim.julian@eawag.ch}

\abbreviations{IR,NMR,UV}

\keywords{American Chemical Society, \LaTeX}

\title[An \textsf{achemso} demo]{Supporting Information. Risk reduction of SARS-CoV-2 transmission
through disinfection of frequently touched surfaces. A Quantitative
Microbial Risk Assessment.}
\makeatletter
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
% pandoc header

\begin{document}
\begin{abstract}

\end{abstract}
\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{concentration-of-virus-in-the-saliva-and-sputum}{%
\subsection{Concentration of virus in the saliva and
sputum}\label{concentration-of-virus-in-the-saliva-and-sputum}}

Viral loads in the saliva or sputum of symptomatic COVID-19 patients
within the first 14 days of symptom onset were used as model input. The
viral loads were measured using quantitative reverse-transcription
polymerase chain reaction (RT-qPCR).\textsuperscript{1--4} The data were
extracted from the figures and is meant to be an approximate number.

PCR data does not distinguish between infectious and non-infectious
virus. No reliable information was found regarding the relationship
between gene copy number and infectivity for
SARS-CoV-2.\textsuperscript{5} Therefore, to convert gene copies to
infections virus, we used a ratio reported for influenza
viruses.\textsuperscript{6}

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Wolfel2020}{}%
(1) Wölfel, R.; Corman, V. M.; Guggemos, W.; Seilmaier, M.; Zange, S.;
Müller, M. A.; Niemeyer, D.; Jones, T. C.; Vollmar, P.; Rothe, C.;
Hoelscher, M.; Bleicker, T.; Brünink, S.; Schneider, J.; Ehmann, R.;
Zwirglmaier, K.; Drosten, C.; Wendtner, C. \emph{Nature} \textbf{2020},
1--14.

\leavevmode\hypertarget{ref-Pan2020}{}%
(2) Pan, Y.; Zhang, D.; Yang, P.; Poon, L. L. M.; Wang, Q. \emph{The
Lancet Infectious Diseases} \textbf{2020}, \emph{20} (4), 411--412.

\leavevmode\hypertarget{ref-Kim2020}{}%
(3) Kim, J. Y.; Ko, J. H.; Kim, Y.; Kim, Y. J.; Kim, J. M.; Chung, Y.
S.; Kim, H. M.; Han, M. G.; Kim, S. Y.; Chin, B. S. \emph{Journal of
Korean Medical Science} \textbf{2020}, \emph{35} (7), 1--7.

\leavevmode\hypertarget{ref-To2020}{}%
(4) To, K. K. W.; Tsang, O. T. Y.; Leung, W. S.; Tam, A. R.; Wu, T. C.;
Lung, D. C.; Yip, C. C. Y.; Cai, J. P.; Chan, J. M. C.; Chik, T. S. H.;
Lau, D. P. L.; Choi, C. Y. C.; Chen, L. L.; Chan, W. M.; Chan, K. H.;
Ip, J. D.; Ng, A. C. K.; Poon, R. W. S.; Luo, C. T.; Cheng, V. C. C.;
Chan, J. F. W.; Hung, I. F. N.; Chen, Z.; Chen, H.; Yuen, K. Y.
\emph{The Lancet Infectious Diseases} \textbf{2020}, \emph{20} (5),
565--574.

\leavevmode\hypertarget{ref-Jefferson2020}{}%
(5) Jefferson, T.; Spencer, E.; Brassey, J.; Heneghan, C. \emph{medRxiv}
\textbf{2020}, 2020.08.04.20167932.

\leavevmode\hypertarget{ref-Ip2016}{}%
(6) Ip, D. K. M.; Lau, L. L. H.; Chan, K. H.; Fang, V. J.; Leung, G. M.;
Peiris, M. J. S.; Cowling, B. J. \emph{Clinical Infectious Diseases}
\textbf{2015}, \emph{62} (4), 431--437.
\end{document}